Taking Dyads Seriously

### 2018-02-19 paper/1_intro.tex
intro edit

We begin with a discussion of these dependencies and an introduction to the AME model. Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the context of dyadic data. Last, to highlight the utility of this approach, we apply the AME model to four recent studies in the international relations (IR) literature. The comparison reveals that in accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables in the literature. Consequently, AME produces results that, at times, differ from those found in the original study as well as from the broader literature. Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies found in the relational data of IR studies. Finally, we show that for each replication our network-based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.

The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest. Thus, the AME approach can be used by scholars in the field to continue to generate substantive insights, while being able to account for the data generating process behind political events of interest. Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.

>>>>

We begin with a discussion of these dependencies and an introduction to the AME model. Next, we conduct a simulation study to show how the AME approach can recover unbiased and well-calibrated regression coefficients in the context of dyadic data. Last, to highlight the utility of this approach, we apply the AME model to three recent studies in the international relations (IR) literature. The comparison reveals that in accounting for observational dependence, AME produces more precise estimates and better-calibrated confidence intervals for key variables in the literature. Consequently, AME produces results that, at times, differ from those found in the original study as well as from the broader literature. Moreover, we demonstrate the latent factor approach offers substantive insights that are often occluded by ignoring the interdependencies found in the relational data of IR studies. Finally, we show that for each replication our network-based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies.

The framework that we present advances statistical analysis of dyadic data by accounting for observational dependence while allowing scholars to test the substantive effect of variables of interest. Thus, the AME approach can be used by scholars in the field to continue to generate substantive insights, while accounting for the data generating process behind political events of interest. Most importantly, the AME approach concentrates on the relational aspect of the field of international relations through a statistical framework that is familiar to most scholars.
